---
title: "Mitt første dokument"
author: "Arnstein"
papersize: a4
editor_options:
  chunk_output_type: console
format: 
  html: 
    paged:
      print: false
  typst: default
  docx: default
editor: visual
---

```{r}
#| label: setup
library(tidyverse)
library(huxtable)
options(
  paged.print = FALSE,
  pillar.print_max = 25, 
  pillar.print_min = 25
  )
```

## Quarto

Min egen tekst.

##### Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code.
You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

1.  første
2.  andre
    c.  punkt
    d.  fffhghg
3.  tredje

```{r}
plot(cars)
```

```{r}
w <- c(2L, 3, "3")
w
```

```{r}
w <- c(2L, 3L, 3)
w
```

```{r}
typeof(w)
```

```{r}
z <- c(3, 5, 1, 9)
```

```{r}
z[-1]
```

```{r}
z[-3]
```

```{r}
z[c(-1, -3)]
```

```{r}
#| eval: false
z[-1, -3]
```

```{r}
z[-c(3:2)]
```

```{r}
z[-c(1, 3)]
```

```{r}
x <- 4:31
```

```{r}
head(x)
```

```{r}
head(x, n = 3)
```

```{r}
head(x, 3)
```

```{r}
#| eval: false
head(3, x)
```

```{r}
head(n = 3, x = x)
```

```{r}
head(n = 3, x = mtcars)
```

```{r}
tail(n = 3, x = mtcars)
```

```{r}
z <- c(3, 5, 1, 9)
```

```{r}
tail(z, n = 2)
```

```{r}
tail(z, n = -2)
```

```{r}
z <- c(z, 7); z
```

```{r}
tail(z, n = 2)
```

### Bruk av tail

```{r}
tail(z, n = -2)
```

```{r}
x <- rnorm(10)
q <- x < 0
```

```{r}
typeof(q)
```

```{r}
q
```

```{r}
q + 5
```

```{r}
l <- list(x = 1:3, y = c("A", letters[4:13]), z = seq(from = 1.3, to = 2.7, by = 0.2))
```

```{r}
ll <- l
```

```{r}
rm(ll)
```

```{r}
#| paged.print: false
mtcars
```

```{r}
typeof(mtcars)
```

```{r}
mtcars$mpg
```

```{r}
typeof(mtcars$mpg)
```

### Hirarki

1.  logical
2.  integer
3.  numeric
4.  character

```{r}
x <- c(TRUE, FALSE, FALSE)
```

```{r}
x
```

```{r}
typeof(x)
```

```{r}
x <- c(x, 1L)
```

```{r}
x
```

```{r}
typeof(x)
```

```{r}
x[5] <- 2.1
```

```{r}
x
```

```{r}
typeof(x)
```

```{r}
x[6] <- "a"
```

```{r}
x
```

```{r}
df <- data.frame(x = 1:3, y = letters[1:3], z = LETTERS[1:3])
```

```{r}
#| eval: false
#| paged.print: false
df[2, c(1, "z")]
```

```{r}
which(row.names(mtcars) == "Toyota Corolla")
```

```{r}
?which
```

```{r}
#| paged.print: false
head(mtcars)
```

```{r}
names(mtcars)
```

```{r}
row.names(mtcars)
```

```{r}
row.names(mtcars) == "Toyota Corolla"
```

```{r}
which(row.names(mtcars) == "Toyota Corolla")
```

```{r}
#| paged.print: false
mtcars[which(row.names(mtcars) == "Toyota Corolla"),]
```

```{r}
#| paged.print: false
mtcars
```

```{r}
#| paged.print: false
mtcars[which(row.names(mtcars) == "Lotus Europa"),]
```

```{r}
#| paged.print: false
mtcars[which(row.names(mtcars) == "Lotus Europa"), 10]
```

```{r}
#| paged.print: false
mtcars[which(row.names(mtcars) == "Lotus Europa"), "gear"]
```

```{r}
modell <- c("Porsche 914-2", "Lotus Europa", "Maserati Bora")
```

```{r}
modell
```

```{r}
row.names(mtcars) %in% modell
```

```{r}
which(row.names(mtcars) %in% modell)
```

```{r}
#| paged.print: false
mtcars[which(row.names(mtcars) %in% modell), ]
```

```{r}
#| paged.print: false
mtcars[which(row.names(mtcars) %in% modell), c("mpg", "hp", "gear")]  
```

```{r}
df <- data.frame(x = 1:3, y = letters[1:3], z = LETTERS[1:3])
```

```{r}
#| paged.print: false
#| echo: true
#| eval: false
df
```

```{r}
df$q <- c(2.1, 0.4, 0.9)
```

## Overskrift

```{r}
df
```

A friend once said:

> It's always better to give than to receive.

```{r}
tib <- as_tibble(df)
```

```{r}
#| paged.print: false
tib
```

```{r}
#| paged.print: false
df
```

```{r}
is.data.frame(tib)
```

```{r}
is.list(tib)
```

```{r}
is.vector(df[2,3])
```

```{r}
is.data.frame(df[2,3])
```

```{r}
is.vector(tib[2,3])
```

```{r}
is.data.frame(tib[2,3])
```

```{r}
w <- tib[2,3]
```

```{r}
#| paged.print: false
w
```

```{r}
df[2,3]
```

```{r}
# define a matirc in R
X <- matrix(c(2, 2, -1, 5, 2, 6), nrow = 3, byrow = FALSE)
```

```{r}
Y <- matrix(c(-3, 2, 2.5, 6, -4, 4), nrow = 2, byrow = FALSE)
```

```{r}
X %*% Y
```

```{r}
Y %*% X
```

```{r}
Y
```

```{r}
X
```

```{r}
#| paged.print: false
subset(cars, subset = cars$speed > 6 & cars$speed < 11)
```

```{r}
cars_sub <- subset(cars, subset = cars$speed > 6 & cars$speed < 11); cars_sub
```

```{r}
#| paged.print: false
cars_sub1 <- subset(cars, subset = cars$speed > 6 & cars$speed < 11 & cars$dist > 15)
```

```{r}
lm1 <- lm(
  formula = dist ~ speed, 
  data = cars
  )
```

```{r}
summary(lm1)
```

```{r}
lm2 <- lm(
  formula = dist ~ speed, 
  data = cars,
  subset = cars$speed > 6 & cars$speed < 11 & cars$dist > 15
  )
```

```{r}
summary(lm2)
```

```{r}
lm3 <- lm(
  formula = dist ~ speed, 
  data = cars_sub1
  )
```

```{r}
summary(lm3)
```

## Ta vare på ting

- `save()`
- `load()`

```{r}
getwd()
```

```{r}
save(cars_sub1, lm1, lm2, file = "my-stuff.RData")
```

```{r}
ls()
```

```{r}
rm(cars_sub1, lm1, lm2)
```

```{r}
ls()
```

```{r}
# Merk at navnene er på plass
load(file = "my-stuff.RData")
```

```{r}
ls()
```

```{r}
huxreg(
        lm1, lm2, lm3,
        error_format = "({statistic})", 
        note         = "{stars}. T statistics in parentheses."
      )
```

```{r}
add_numbers <- function(x, y) {
  x + y
}
```

```{r}
add_numbers(3, 4)
```

```{r}
add_numbers <- function(x, y) {
  res = x + y
  res
}
```

```{r}
add_numbers(3, 4)
```

```{r}
add_numbers <- function(x = 0, y = 0) {
  res = x + y
  res
}
```

```{r}
add_numbers()
```

```{r}
add_numbers(5)
```

```{r}
df
```

```{r}
start_stupid_fun <- function(x1 = 0, n = 10, d = 1) {
# initialise x
x = vector(mode = "numeric", length = n)
x[1] = x1

for (i in seq_along(x[-1])) {
  x[i + 1] = x[i] + d
}
print(x)
}
```

```{r}
start_stupid_fun()
```

```{r}
start_stupid_fun(x1 = 17, n = 6)
```

```{r}
start_stupid_fun(x1 = 17, n = 6, d = 2)
```

```{r}
start_stupid_fun(x1 = 17, n = 6, d = 0.5)
```

```{r}
head_and_tail <- function(x) {
  c(head(x), tail(x))
}
```

```{r}
head_and_tail(x = mtcars)
```

```{r}
mtcars
```

```{r}
head_and_tail <- function(x, n=3L) {
n_vars = length(names(x))
res = rbind(
        rep('Head', n_vars),
        head(x, n = n),
        rep('*-*-*', n_vars),
        rep('Tail', n_vars),
        tail(x, n = n)
        )
as_tibble(res)
}
```

```{r}
head_and_tail(x = mtcars, n = 2)
```

```{r}
head_and_tail(letters, n = 2)
```



